### 2018-11-29 Sections/state_of_the_art.tex
Editing SOTA

In this section we present a brief history of recent major developments and the state-of-the-art of some of the most important quantum technologies and protocols relevant to the quantum internet, with a view to understand trends in the development of the field, so as to position us to gauge the rate of progress and make educated predictions about future development. With this we aim to shed light on what quantum networking services could be readily implemented today, using present-day technology, and what is likely to be viable in the near and more distant future.

@@

The reader disinterested in a history lesson might comfortably skip this section, without serious risk of missing out on explanations of important topics (no new conceptual material is introduced in this part).

Tables summarising timelines of major developments and their referencing are presented in bulk at the end of this part in Sec.~\ref{sec:timeline_tables}.

@@

\section{Entanglement swapping \& quantum repeaters} \index{Entanglement!Swapping} \index{Quantum repeater networks}

\dropcap{C}{losely} related to entanglement distribution is entanglement swapping (Sec.~\ref{sec:swapping}), where the goal is to entangle two remote parties, each of whom have one half of two distinct Bell pairs. As discussed in Sec.~\ref{sec:ent_ultimate}, entanglement swapping will be of fundamental importance in networks treating entanglement distribution as the most fundamental elementary resource, making it of great importance to quantum networking, allowing the distribution of entanglement between distant parties, who are unable to communicate directly (i.e the entanglement distribution is `mediated').

Tabs.~\ref{tab:entanglement_swap1} \& \ref{tab:entanglement_swap2} summarise major recent developments in entanglement swapping and quantum repeater networks (in which entanglement swapping is a basic primitive).

@@

Tab.~\ref{tab:QKD_table} summarises some major milestones in the development of the field.

@@

In the context of the quantum internet, the most important class of quantum states is of course the optical ones, since light will almost inevitably be the primary information carrier in such a network.

@@

Coherent states are well approximated by laser or maser light. Nowadays thousands of types of lasers are known with different power, temporal, spatial and spectral parameters, and the technology is already very mature and available commercially. Here we only introduce some of the basic concepts of lasers related to quantum networking.

A laser can be classified as operating in either a continuous or pulsed regime. Most laser diodes used in communication systems are continuous. But they can also be externally carved at some rate by modulators to create pulsed light. Usually, pulsed lasers are created by the technique of Q-switching or mode-locking.

Different applications require lasers with different output power. Typical output powers of single-mode laser diodes are some tens of milliwatts, up to at most a few hundred of mW. However, multiple transverse mode diode lasers can reach up to some tens of Watts, and can be used as pump sources for high-quality and high-power single-mode solid-state lasers. Such single mode diode-pumped solid-state lasers can be further mode-locked to output femtosecond pulses, reaching as far as an average power of tens of Watts. Pulsed lasers can also be characterised with the peak power of each pulse. The peak power of a pulsed laser is many orders of magnitude greater than its average power.

@@

A highly attenuated laser can be used as a good approximation to a single-photon source when no more than one single-photon is used in an interferometry experiment, such as QKD. Otherwise, heralded or deterministic single-photon sources are required.

Tab.~\ref{tab:single_photon_state} summarises some of the notable developments. A good review about solid-state single-photon emitters is \cite{bib:aharonovich2016solid}. Different sources are compared in the review by \cite{bib:eisaman2011}.

@@

\section{Non-optical systems} \index{Matter qubits}

\dropcap{A}{lthough} light is likely to be the information \textit{carrier} of the future quantum internet, it may very well not win the race of becoming the preferred means for computation or storage. Here, there are countless other possibilities, of which we now outline a few.

@@

A variety of techniques have been developed to create squeezing and entanglement in atomic ensembles. The main methods exploit atom-light interactions in cold gases, or interaction between particles such as atom-atom collisions in Bose-Einstein condensates, or combined electrostatic and ion-light interaction in ion chains. Atom-light interaction currently represents the most mature method, and exhibits the highest squeezing of 20.1dB via an optical-cavity-based measurement \cite{bib:hosten2016measurement}.

@@

Many experimental methods have been developed for measuring and manipulating individual quantum systems, including single atoms in a cavity, trapped ions, and neutral atoms in an optical lattice.

@@

In quantum dots, the recombination of an electron-hole pair, known as an \textit{exciton}, emits a single photon. Photon sources based on this concept are rapidly gaining popularity in present-day labs, and a summary of developments in this area is provided in Tab.~\ref{tab:quantum_dots}.

@@

An nitrogen-vacancy (NV) centres in diamond refers to a nitrogen (N) atom replacing a carbon atom and neighbouring one vacancy (V) \cite{bib:doherty2013nitrogen}. In such centres, both electron and nuclear spins can exhibit long coherence times ($>$1ms for the electron spin, and $>$1s for the nuclear spin) even at room temperature \cite{bib:balasubramanian2009ultralong, bib:neumann2010quantum, bib:maurer2012room}.

Tab.~\ref{tab:NV_centres} summarises some of the recent developments in NV centre technology. Further material can be found in several review papers \cite{bib:doherty2013nitrogen, bib:atature2018material, bib:awschalom2018quantum}.

@@

Tab.~\ref{tab:superconducting} summarises some of the recent achievements in superconducting quantum technology. The review paper \cite{bib:xiang2013hybrid} provides excellent follow-up material for more details.

@@

\dropcap{M}{any} quantum information protocols have been developing as two versions in parallel -- a discrete-variable version, and a continuous-variable version.

@@

There are a variety of single-photon detectors, detailed description of which can be found in several excellent review papers \cite{bib:eisaman2011, bib:hadfield2009}.

@@

\subsection{Homodyning}

Tab.~\ref{tab:homodyning} summarises some of the notable developments in homodyne detection, the most fundamental measurement device in CV optical systems.

@@

\dropcap{Q}{uantum} states can be encoded in various degrees of freedom of a single photon. Tab.~\ref{tab:evolutionofstates} summarises some of the major developments in the controlled evolution of optical states. There are two primary types of optical circuits: optical waveguides, and fibre loops.

Optical waveguides can be integrated onto a small chip, which is alignment-free, compact, and phase-stable. Fibre-loop schemes are extremely resource-frugal -- resource requirements are essentially constant, and do not grow with the complexity of the interferometer being implemented. Some of these state-of-the-art techniques are summarised in Tab.~\ref{tab:waveguide_fibre}.

@@

\dropcap{N}{umerous} excellent review articles exist \cite{bib:lvovsky2009optical, bib:simon2010quantum, bib:sangouard2011quantum, bib:bussieres2013prospective, bib:reiserer2015cavity}, providing detailed explanations of quantum memory architectures. Quantum memory has such widespread utility that it has become a major research area in its own right.

The simplest approach to storing light is using an optical delay line, such as an optical fibre. Aside from loss, these also suffer that the storage time is fixed by the delay length, making on-demand output extremely challenging without first having access to high-speed ($\sim$GHz) optical switches.

Alternatively, light can be stored in a high-Q cavity. The light cycles back and forth between the reflecting boundaries, allowing it to be injected into and retrieved from the cavity \cite{bib:pittman2002single, bib:pittman2002cyclical, bib:leung2006quantum, bib:maitre1997quantum, bib:tanabe2007trapping, bib:tanabe2009dynamic}. Unfortunately, the storage of light in cavities suffers from a tradeoff between short cycle time and long storage time, which limits the efficiency. Therefore, whereas optical delay lines and nano-cavities could be appropriate for obtaining on-demand single photons from heralded sources \cite{bib:saglamyurek2015quantum, bib:jin2015telecom}, they may not be suitable for quantum memory or quantum repeaters.

Tab.~\ref{tab:memory} summarises some of the notable developments in quantum memory.

@@

\dropcap{T}{he} Holy Grail\index{Holy Grail} of scalable, universal quantum computation -- perhaps the most exciting of future quantum developments -- has been developing rapidly across various physical platforms simultaneously in parallel, including: trapped atoms/ions; nuclear magnetic resonance (NMR); photons; superconductors; spins in silicon; and many, many more. An excellent point of reference is the review by \cite{bib:ladd2010quantum}.

Tab.~\ref{tab:quantumcomputer} summarises some of current records for the number of qubits and attained fidelities in different physical systems for implementing quantum computation. Most noteworthy, IBM and Google recently announced their record-breaking 50-qubit and 72-qubit superconducting quantum processors respectively, a key step towards achieving the coveted goal of quantum supremacy \cite{bib:savage2018quantum, bib:neill2018blueprint, bib:harrow2017quantum}.

@@

\section{Timelines \& references}\label{sec:timeline_tables}

\dropcap{T}{he} development timelines of the technologies referred to in this part, and their references, now follow in table form.

@@

		The first experimental demonstration of entanglement swapping. By pumping a BBO crystal in a double-pass configuration, two pairs of polarization-entangled photons were generated to demonstrate the scheme. A visibility of \mbox{$0.65$} was observed, which clearly surpasses the classical limit of \mbox{$0.5$}. This was later improved in 2001 to a visibility of \mbox{$0.84$} , which violates the Bell inequality (for which the threshold is $0.71$).  & \cite{bib:PRL_80_3891, bib:jennewein2001experimental} \\

@@

		Delayed-choice entanglement swapping was designed and realized. This was performed by adding two 10m optical fibre delays of about 50ns for both outputs of the Bell state measurement. & \cite{bib:PRL_88_017903}\\

@@

		A delayed-choice entanglement swapping experiment with vacuum-one-photon quantum states was realized. 

@@

		By designing a special interferometer to realize active switching between Bell state measurement and separable state measurement, an entanglement swapping experiment with active delayed-choice was demonstrated. 

@@

		By using three pairs of polarization-entangled photons and conducting two Bell state measurements, a multistage entanglement swapping was realized.&\cite{bib:goebel08}\\

@@

		Hybrid entanglement swapping between discrete-variables and Continuous variables optical systems was realized experimentally.&\cite{bib:takeda2015entanglement}\\

@@

		 Entanglement swapping over 100 km optical fiber with independent entangled photon-pair sources.&\cite{bib:sun2017entanglement}\\

@@

	Quantum cryptography using polarized photons in optical fibre over more than 1km. & \cite{bib:EL_23_383} \\

@@

	Decoy-state QKD over a 200km optical fibre cable through photon polarization with a final key rate of 15Hz & \cite{bib:OptExp_18_8587} \\

@@

	First realization of a differential phase-shift (DPS) QKD protocol over a 42.1dB lossy channel and 200km of optical dispersion-shifted fibre. & \cite{bib:NP_1_343} \\

@@

	The DPS protocol over 50dB channel loss and 260km of optical fibre using superconductive detectors was realized. This is the first implementation of QKD over more than 50dB channel loss.&\cite{bib:OL_37_1008}\\

@@

	The coherent one way (COW) protocol QKD system with a maximum range of 250km at 42.6dB channel loss using ultra-low-loss fibre, with secret bit rates up to 15Hz was realized.&\cite{bib:NJP_11_075003}\\

@@

Since 2006, the EU has established a `SECOQ' network, combining the efforts of 41 research and industrial organizations from 12 countries, including the UK, France, Germany, and Austria. A typical network employing a trusted repeater paradigm, with 6 nodes and 8 links was demonstrated in Vienna. & \cite{bib:NJP_11_075001}\\

@@

 A metropolitan all-pass and intercity quantum communication network in field fiber for four nodes was designed. Any two nodes can be connected in the network, QKD between arbitrary pairs of users.&\cite{bib:OpEx_18_27217}\\

@@

A Brazilian group demonstrated the MDI QKD protocol using a polarization encoding scheme.&\cite{bib:PRA_88_052303}\\

@@

\begin{table*}[!htbp]
	\begin{tabular}{|p{0.755\linewidth}|p{0.22\linewidth}|}
		\hline
		Summary & References \& years \\
		\hline \hline
        In 1999, the world's first multi-particle entanglement -- 3-photon GHZ entanglement was generated. After that, Pan and his colleagues broke the records continuously, realising 4-, 5-, 6-, 8-, 10-, 12-photon entanglement and 10- and 18-qubit hyper-entanglement with multiple degrees of freedom. & \cite{bib:bouwmeester1999observation,  bib:wang201818, bib:zhong201812}  \\
		\hline
		8-photon optical cluster states have been realised, and was used to demonstrate topological error correction against a single error on any qubit. &  \cite{bib:yao2012experimental}  \\
		\hline
		5-photon NOON states have been realised by mixing quantum and classical light. & \cite{bib:afek2010high} \\
		\hline
		6-photon Holland-Burnett states have been realised both at visible wavelengths and at telecom wavelengths. & \cite{bib:xiang2012optimal,  bib:jin2016detection} \\
		\hline
		Entangled photon pairs were created with dimension up to \mbox{$15\times 15$} on a programmable silicon integrated circuit. & \cite{bib:wang2018multidimensional} \\
		\hline
		Zeilinger and his colleagues created orbital angular momentum (OAM) entangled photon pairs differing by 600 in quantum number. Four yeas later, they created a ($100\times 100$)-dimensional OAM entangled photon pairs. & \cite{bib:fickler2012quantum} \\
		\hline
		Furusawa \textit{et al.} deterministically generate and fully characterize a continuous-variable cluster state containing more than 10,000 entangled modes. Three years later, they improved the system to more than one million modes. & \cite{bib:yokoyama2013ultra, bib:yoshikawa2016invited} \\
		\hline
        Cat states $\propto \left| \alpha  \right\rangle  - \left| { - \alpha } \right\rangle $ with small ``size" ${\left| \alpha  \right|^2} \lesssim 1$ were generated by reflecting a small fraction of squeezed vacuum state toward a single photon avalanche diodes. & \cite{bib:neergaard2006generation,  bib:ourjoumtsev2006generating,  bib:wakui2007photon} \\
        \hline
        Large ``size" cat states (e.g ${\left| \alpha  \right|^2} > 2.3$) were prepared by reflecting a half of photon number state $\left| n \right\rangle$ toward a momentum quadrature homodyne detector. & \cite{bib:ourjoumtsev2007generation, bib:takahashi2008generation} \\
		\hline
		Remote entanglement of two independent cat states was created by interfering small fractions of each pulse. & \cite{bib:ourjoumtsev2009preparation} \\
		\hline
	\end{tabular}
	\captionspacetab \caption{Developments in entangled state preparation, based on non-linear optics.} \label{tab:entangled_states}
\end{table*}

\begin{table*}[!htbp]
	\begin{tabular}{|p{0.755\linewidth}|p{0.22\linewidth}|}
		\hline
		Summary & References \& years \\
		\hline \hline
		Atom-light interaction currently represents the most mature method to creat atomic squeezing, and owns the highest squeezing of 20.1 dB via an optical-cavity-based measurement. & \cite{bib:hosten2016measurement} \\
		\hline
		The detection of a single photon prepares almost 3000 atoms to an entangled Dicke state. The state was reconstructed with a negative-valued Wigner function --- an important hallmark of non-classicality. &  \cite{bib:mcconnell2015entanglement} \\
		\hline
		In 2010, two ensembles was entangled by storage of two entangled light fields; and ten years later, four quantum memories was entangled via spin wave. & \cite{bib:lukin2000entanglement, bib:choi2010entanglement} \\
		\hline
		Bose-Einstein condensate was used to create large ensembles of up to ${10^4}$ pair-correlated atoms with an interferometric sensitivity $-1.61$ decibels beyond the shot noise limit. & \cite{bib:lucke2011twin} \\
		\hline
	\end{tabular}
	\captionspacetab \caption{Notable developments in atomic ensembles.} \label{tab:atomic_ensembles}
\end{table*}

\begin{table*}[!htbp]
	\begin{tabular}{|p{0.755\linewidth}|p{0.22\linewidth}|}
		\hline
		Summary & References \& years \\
		\hline \hline		
		In 2000, Haroche and his colleagues generated entanglement among two atoms and a single-photon cavity mode. Later, entanglement was demonstrated between two photons sequentially emitted by the same single atom in a cavity. The entanglement fidelity of 86\% between the two photons was obtained, and importantly, the entanglement-generation efficiency is 15\%, a drastic improvement compared to free-space experiments with single atoms. & \cite{bib:rauschenbeutel2000step, bib:wilk2007single, bib:blinov2004observation} \\
		\hline
		In the atom-cavity system, one of the largest Schr{\"o}dinger cat state was created with spin of 25 on a Rydberg atom. & \cite{bib:facon2016sensitive} \\
		\hline
		In the recent year, Blatt and his colleagues reported scalable and deterministic generation of multi-ions entanglement, including 8-qubit W state and 14-qubit GHZ state. & \cite{bib:haffner2005scalable, bib:monz2011}\\
		\hline
		It is up to 50 qubits for the record of arbitrary coherent addressing of individual neutral atoms in a 3D optical lattice. & \cite{bib:wang2015coherent} \\
		\hline		
		Several groups reported generation and detection of two neural-atom entanglement in optical lattices with single-site-resolved technique . & \cite{bib:kaufman2015entangling, bib:islam2015measuring, bib:dai2016generation}\\
		\hline
	\end{tabular}
	\captionspacetab \caption{Notable developments in single-atom technology.} \label{tab:single_atoms}
\end{table*}

>>>>

In this section we present a brief history of recent major developments and the state-of-the-art of some of the most important quantum technologies and protocols relevant to the quantum internet, with a view to understand trends in the development of the field, so as to position us to gauge the rate of progress and make educated predictions about future developments -- the timing of future developments is likely to have big strategic implications, and it is important that we are able to anticipate technological transformations as radical as the quantum ones, so that we may be prepared for their fallout and consequences.

With this we aim to shed light on what quantum networking services could be readily implemented today, using present-day technology, and what is likely to be viable in the near and more distant future.

@@

The reader disinterested in a history lesson might comfortably skip this part, without serious risk of missing out on explanations of important topics (no new conceptual material is introduced in this part).

The tables summarising chronological timelines of major developments, and their referencing, are presented in bulk at the end of this part in Sec.~\ref{sec:timeline_tables}.

@@

\section{Entanglement swapping \& quantum repeaters} \index{Entanglement!Swapping} \index{Quantum repeater networks}
	\index{Entanglement!Distribution}

\dropcap{C}{losely} related to entanglement distribution is entanglement swapping (Sec.~\ref{sec:swapping}), where the goal is to entangle two remote parties, each of whom have one half of two distinct Bell pairs. As discussed in Sec.~\ref{sec:ent_ultimate}, entanglement swapping will be of fundamental importance in networks treating entanglement distribution as the most fundamental elementary resource, making it of great importance to quantum networking, allowing the distribution of entanglement between distant parties, who are unable to communicate directly (i.e indirect entanglement distribution is `mediated' by the network, in the absence of any P2P channels\index{P2P channel} between end users).

Tabs.~\ref{tab:entanglement_swap1} \& \ref{tab:entanglement_swap2} summarise major recent developments in entanglement swapping and quantum repeater networks (in which entanglement swapping is also a basic primitive).

@@

Tab.~\ref{tab:QKD_table} summarises some major milestones in the development and deployment of QKD.

@@

In the context of the quantum internet, the most important class of quantum states is of course the optical ones, since light will almost inevitably be the primary information carrier in such a future network.

@@

Coherent states are well approximated by laser\index{Lasers} or maser\index{Masers} light. Nowadays thousands of types of lasers are known with different power, temporal, spatial and spectral parameters, and the technology is already very mature and available commercially. Here we only introduce some of the basic concepts of lasers related to quantum networking.

A laser can be classified as operating in either a continuous\index{Continuous wave lasers} or pulsed\index{Pulsed lasers} regime. Most laser diodes\index{Lasers!Diodes} used in communication systems are continuous. But they can also be externally carved at some rate by modulators to create pulsed light. Usually, pulsed lasers are created by the technique of Q-switching\index{Q-switching} or mode-locking\index{Mode-locking}.

Different applications require lasers with different output power. Typical output powers of single-mode laser diodes are some tens of milliwatts, up to at most a few hundred mW. However, multiple transverse mode diode lasers\index{Transverse mode diode lasers} can reach up to some tens of Watts, and can be used as pump sources for high-quality and high-power single-mode solid-state lasers. Such single-mode diode-pumped solid-state lasers can be further mode-locked to output femtosecond pulses, reaching as far as an average power of tens of Watts. Pulsed lasers can also be characterised by the peak power of each pulse. The peak power\index{Peak power} of a pulsed laser is many orders of magnitude greater than its average power\index{Average power}.

@@

A highly attenuated laser can be used as a good approximation to a single-photon source when no more than one single-photon is used in an interferometric experiment, such as QKD. Otherwise, heralded or deterministic single-photon sources are required. This is because while the photo-statistics of a single-photon mimic those of a laser, the photo-statistics of multi-photon states do not, irrespective of the laser's power -- coherent light and multi-photon states obey fundamentally different rules of statistics.

Tab.~\ref{tab:single_photon_state} summarises some of the notable developments. A good review of solid-state single-photon emitters is \cite{bib:aharonovich2016solid}. Different sources are compared in the review by \cite{bib:eisaman2011}.

@@

\section{Non-optical systems}\index{Non-optical systems}\index{Matter qubits}

\dropcap{A}{lthough} light is likely to be the information \textit{carrier} of the future quantum internet, it may very well not win the race of becoming the preferred means for computation or storage. Here, there are countless other possibilities to consider, of which we now outline a few.

@@

A variety of techniques have been developed to create squeezing\index{Squeezing!Atomic ensembles} and entanglement in atomic ensembles. The main methods exploit atom-light interactions in cold gases, or interactions between particles such as atom-atom collisions in Bose-Einstein condensates\index{Bose-Einstein condensates}, or combined electrostatic and ion-light interactions in ion chains. Atom-light interactions currently represent the most mature technology, and exhibit the highest squeezing of 20.1dB via an optical-cavity-based measurement \cite{bib:hosten2016measurement}.

@@

Many experimental methods have been developed for measuring and manipulating individual quantum systems, including single atoms in a cavity, trapped ions\index{Trapped ions}, and neutral atoms\index{Neutral atoms} in an optical lattice\index{Optical lattice}.

@@

In quantum dots, the recombination of an electron-hole pair\index{Electron-hole pair}, known as an \textit{exciton}\index{Exciton}, emits a single photon. Photon sources based on this concept are rapidly gaining popularity in present-day labs, and a summary of developments in this area is provided in Tab.~\ref{tab:quantum_dots}.

@@

A nitrogen-vacancy (NV) centre in diamond refers to a nitrogen (N) atom replacing a carbon atom in the crystal lattice, neighbouring a single vacancy (V) defect \cite{bib:doherty2013nitrogen}. In such centres, both electrons and nuclear spins can exhibit long coherence times ($>$1ms for the electron spin, and $>$1s for the nuclear spin) even at room temperature \cite{bib:balasubramanian2009ultralong, bib:neumann2010quantum, bib:maurer2012room}.

Tab.~\ref{tab:NV_centres} summarises some of the recent developments in NV centre technology. Further material can be found in the review papers \cite{bib:doherty2013nitrogen, bib:atature2018material, bib:awschalom2018quantum}.

@@

Tab.~\ref{tab:superconducting} summarises some of the recent achievements in superconducting quantum technology. The review paper \cite{bib:xiang2013hybrid} provides excellent follow-up material for further details.

@@

\dropcap{M}{any} quantum information protocols have been developing as two versions in parallel -- a discrete-variable\index{Discrete-variables} version, and a continuous-variable\index{Continuous-variables} version.

@@

There are a variety of single-photon detectors, detailed descriptions of which can be found in several excellent review papers \cite{bib:eisaman2011, bib:hadfield2009}.

@@

\subsection{Homodyning}\index{Homodyne detection}

Tab.~\ref{tab:homodyning} summarises some of the notable developments in homodyne detection, the most fundamental approach to quantum measurement in CV optical systems.

@@

\dropcap{Q}{uantum} states can be encoded in various degrees of freedom of a single photon. Tab.~\ref{tab:evolutionofstates} summarises some of the major developments in the controlled evolution of optical states.

There are two primary types of optical circuits: optical waveguides\index{Optical waveguides}, and fibre-loops\index{Fibre-loops}.

Optical waveguides can be miniaturised and integrated onto a small chip, which is alignment-free, compact, and phase-stable\index{Phase-stability}.

Fibre-loop schemes are extremely resource-frugal -- resource requirements are essentially constant, and do not grow with the complexity of the underlying interferometer being implemented.

Some of these techniques are summarised in Tab.~\ref{tab:waveguide_fibre}.

@@

\dropcap{N}{umerous} excellent review articles exist \cite{bib:lvovsky2009optical, bib:simon2010quantum, bib:sangouard2011quantum, bib:bussieres2013prospective, bib:reiserer2015cavity}, providing detailed explanations of quantum memory schemes. Quantum memory is so ubiquitous and has such widespread utility that it has become a major research area in its own right.

The simplest approach to storing light is using an optical delay line\index{Delay line}, such as an optical fibre. Aside from loss, these also suffer that the storage time is fixed by the delay length, making on-demand output\index{On-demand output} extremely challenging without first having access to high-speed ($\sim$GHz) optical switches\index{Optical!Switches}.

Alternatively, light can be stored in a high-Q cavity\index{Cavities}. The light cycles back and forth between the reflecting boundaries, allowing it to be injected into and retrieved from the cavity \cite{bib:pittman2002single, bib:pittman2002cyclical, bib:leung2006quantum, bib:maitre1997quantum, bib:tanabe2007trapping, bib:tanabe2009dynamic}. Unfortunately, the storage of light in cavities suffers from a tradeoff between short cycle time\index{Cycle time} and long storage time\index{Storage time}, which limits the efficiency. Therefore, whereas optical delay lines and nano-cavities could be appropriate for obtaining on-demand single photons from heralded sources \cite{bib:saglamyurek2015quantum, bib:jin2015telecom}, they may not be suitable for quantum memory or quantum repeaters.

Tab.~\ref{tab:memory} summarises some of the notable developments in quantum memories.

@@

\dropcap{T}{he} Holy Grail\index{Holy Grail} of scalable, universal quantum computation -- perhaps the most exciting of future quantum developments -- has been developing rapidly across various physical platforms simultaneously in parallel, including: trapped atoms/ions\index{Trapped atoms}; nuclear magnetic resonance (NMR)\index{Nuclear magnetic resonance (NMR)}; photons; superconductors; spins in silicon\index{Spins in silicon}; and many, many more. An excellent point of reference is the review by \cite{bib:ladd2010quantum}.

Tab.~\ref{tab:quantumcomputer} summarises some of current records for the number of qubits and attained fidelities in different physical systems for implementing quantum computation. Most noteworthy, IBM\index{IBM} and Google\index{Google} recently announced their record-breaking 50- and 72-qubit superconducting quantum processors respectively, a key step towards achieving the coveted goal of quantum supremacy\index{Quantum supremacy} \cite{bib:savage2018quantum, bib:neill2018blueprint, bib:harrow2017quantum}.

@@

\section{Timelines \& references}\index{Timelines}\label{sec:timeline_tables}

\dropcap{T}{he} historical development timelines of the technologies referred to earlier in this part, and their references, now follow in table form.

@@

		The first experimental demonstration of entanglement swapping. By pumping a BBO crystal in a double-pass configuration, two pairs of polarisation-entangled photons were generated to demonstrate the scheme. A visibility of \mbox{$0.65$} was observed, which clearly surpasses the classical limit of \mbox{$0.5$}. This was later improved in 2001 to a visibility of \mbox{$0.84$} , which violates the Bell inequality (for which the threshold is $0.71$).  & \cite{bib:PRL_80_3891, bib:jennewein2001experimental} \\

@@

		Delayed-choice entanglement swapping was designed and realised. This was performed by adding two 10m optical fibre delays of about 50ns for both outputs of the Bell state measurement. & \cite{bib:PRL_88_017903}\\

@@

		A delayed-choice entanglement swapping experiment with vacuum-one-photon quantum states was realised. 

@@

		By designing a special interferometer to realise active switching between Bell state measurement and separable state measurement, an entanglement swapping experiment with active delayed-choice was demonstrated. 

@@

		By using three pairs of polarisation-entangled photons and conducting two Bell state measurements, a multistage entanglement swapping was realised.&\cite{bib:goebel08}\\

@@

		Hybrid entanglement swapping between discrete-variables and Continuous variables optical systems was realised experimentally.&\cite{bib:takeda2015entanglement}\\

@@

		 Entanglement swapping over 100 km optical fibre with independent entangled photon-pair sources.&\cite{bib:sun2017entanglement}\\

@@

	Quantum cryptography using polarised photons in optical fibre over more than 1km. & \cite{bib:EL_23_383} \\

@@

	Decoy-state QKD over a 200km optical fibre cable through photon polarisation with a final key rate of 15Hz & \cite{bib:OptExp_18_8587} \\

@@

	First realisation of a differential phase-shift (DPS) QKD protocol over a 42.1dB lossy channel and 200km of optical dispersion-shifted fibre. & \cite{bib:NP_1_343} \\

@@

	The DPS protocol over 50dB channel loss and 260km of optical fibre using superconductive detectors was realised. This is the first implementation of QKD over more than 50dB channel loss.&\cite{bib:OL_37_1008}\\

@@

	The coherent one way (COW) protocol QKD system with a maximum range of 250km at 42.6dB channel loss using ultra-low-loss fibre, with secret bit rates up to 15Hz was realised.&\cite{bib:NJP_11_075003}\\

@@

Since 2006, the EU has established a `SECOQ' network, combining the efforts of 41 research and industrial organisations from 12 countries, including the UK, France, Germany, and Austria. A typical network employing a trusted repeater paradigm, with 6 nodes and 8 links was demonstrated in Vienna. & \cite{bib:NJP_11_075001}\\

@@

 A metropolitan all-pass and intercity quantum communication network in field fibre for four nodes was designed. Any two nodes can be connected in the network, QKD between arbitrary pairs of users.&\cite{bib:OpEx_18_27217}\\

@@

A Brazilian group demonstrated the MDI QKD protocol using a polarisation encoding scheme.&\cite{bib:PRA_88_052303}\\

@@

\begin{table*}[!htbp]
	\begin{tabular}{|p{0.755\linewidth}|p{0.22\linewidth}|}
		\hline
		Summary & References \& years \\
		\hline \hline
        In 1999, the world's first multi-particle entanglement -- 3-photon GHZ entanglement was generated. After that, Pan and his colleagues broke the records continuously, realising 4-, 5-, 6-, 8-, 10-, 12-photon entanglement and 10- and 18-qubit hyper-entanglement with multiple degrees of freedom. & \cite{bib:bouwmeester1999observation, bib:wang201818, bib:zhong201812}  \\
		\hline
		8-photon optical cluster states have been realised, and was used to demonstrate topological error correction against a single error on any qubit. &  \cite{bib:yao2012experimental}  \\
		\hline
		5-photon NOON states have been realised by mixing quantum and classical light. & \cite{bib:afek2010high} \\
		\hline
		6-photon Holland-Burnett states have been realised both at visible wavelengths and at telecom wavelengths. & \cite{bib:xiang2012optimal,  bib:jin2016detection} \\
		\hline
		Entangled photon pairs were created with dimension up to \mbox{$15\times 15$} on a programmable silicon integrated circuit. & \cite{bib:wang2018multidimensional} \\
		\hline
		Zeilinger and his colleagues created orbital angular momentum (OAM) entangled photon pairs differing by 600 in quantum number. Four yeas later, they created a ($100\times 100$)-dimensional OAM entangled photon pairs. & \cite{bib:fickler2012quantum} \\
		\hline
		Furusawa \textit{et al.} deterministically generate and fully characterise a continuous-variable cluster state containing more than 10,000 entangled modes. Three years later, they improved the system to more than one million modes. & \cite{bib:yokoyama2013ultra, bib:yoshikawa2016invited} \\
		\hline
        Cat states $\propto \left| \alpha  \right\rangle  - \left| { - \alpha } \right\rangle $ with small ``size" ${\left| \alpha  \right|^2} \lesssim 1$ were generated by reflecting a small fraction of squeezed vacuum state toward a single photon avalanche diodes. & \cite{bib:neergaard2006generation,  bib:ourjoumtsev2006generating, bib:wakui2007photon} \\
        \hline
        Large ``size" cat states (e.g ${\left| \alpha  \right|^2} > 2.3$) were prepared by reflecting a half of photon number state $\left| n \right\rangle$ toward a momentum quadrature homodyne detector. & \cite{bib:ourjoumtsev2007generation, bib:takahashi2008generation} \\
		\hline
		Remote entanglement of two independent cat states was created by interfering small fractions of each pulse. & \cite{bib:ourjoumtsev2009preparation} \\
		\hline
	\end{tabular}
	\captionspacetab \caption{Developments in entangled state preparation, based on non-linear optics.} \label{tab:entangled_states}
\end{table*}

\begin{table*}[!htbp]
	\begin{tabular}{|p{0.755\linewidth}|p{0.22\linewidth}|}
		\hline
		Summary & References \& years \\
		\hline \hline
		Atom-light interaction currently represents the most mature method to create atomic squeezing, and owns the highest squeezing of 20.1 dB via an optical-cavity-based measurement. & \cite{bib:hosten2016measurement} \\
		\hline
		The detection of a single photon prepares almost 3000 atoms to an entangled Dicke state. The state was reconstructed with a negative-valued Wigner function --- an important hallmark of non-classicality. &  \cite{bib:mcconnell2015entanglement} \\
		\hline
		In 2010, two ensembles was entangled by storage of two entangled light fields; and ten years later, four quantum memories was entangled via spin wave. & \cite{bib:lukin2000entanglement, bib:choi2010entanglement} \\
		\hline
		Bose-Einstein condensate was used to create large ensembles of up to ${10^4}$ pair-correlated atoms with an interferometric sensitivity $-1.61$ decibels beyond the shot noise limit. & \cite{bib:lucke2011twin} \\
		\hline
	\end{tabular}
	\captionspacetab \caption{Notable developments in atomic ensembles.} \label{tab:atomic_ensembles}
\end{table*}

\begin{table*}[!htbp]
	\begin{tabular}{|p{0.755\linewidth}|p{0.22\linewidth}|}
		\hline
		Summary & References \& years \\
		\hline \hline		
		In 2000, Haroche and his colleagues generated entanglement among two atoms and a single-photon cavity mode. Later, entanglement was demonstrated between two photons sequentially emitted by the same single atom in a cavity. The entanglement fidelity of 86\% between the two photons was obtained, and importantly, the entanglement-generation efficiency is 15\%, a drastic improvement compared to free-space experiments with single atoms. & \cite{bib:rauschenbeutel2000step, bib:wilk2007single, bib:blinov2004observation} \\
		\hline
		In the atom-cavity system, one of the largest Schr{\"o}dinger cat state was created with spin of 25 on a Rydberg atom. & \cite{bib:facon2016sensitive} \\
		\hline
		In the recent year, Blatt and his colleagues reported scalable and deterministic generation of multi-ions entanglement, including 8-qubit W-state and 14-qubit GHZ state. & \cite{bib:haffner2005scalable, bib:monz2011}\\
		\hline
		It is up to 50 qubits for the record of arbitrary coherent addressing of individual neutral atoms in a 3D optical lattice. & \cite{bib:wang2015coherent} \\
		\hline		
		Several groups reported generation and detection of two neural-atom entanglement in optical lattices with single-site-resolved technique . & \cite{bib:kaufman2015entangling, bib:islam2015measuring, bib:dai2016generation}\\
		\hline
	\end{tabular}
	\captionspacetab \caption{Notable developments in single-atom technology.} \label{tab:single_atoms}
\end{table*}
